%Input preamble, math commands, environments, etc. from a saingle file. 
\input{HelperFiles/BenPorathcopy}
\begin{document}

\title{Human Capital Accumulation and Earnings Dynamics over the Life Cycle: Lessons from the Ben-Porath Model}
\author{Jorge Luis Garc\'{i}a\thanks{Department of Economics, The University of Chicago (jorgelgarcia@uchicago.edu).} \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \
 \ \ \ \ \ \ \ \ \ James J. Heckman\thanks{The University of Chicago; American Bar Foundation; University College Dublin (jjh@uchicago.edu).} $^{,}$\thanks{We thank  Will Burgo for outstanding research assistance.}}
\date{This Draft: \today}
\maketitle

%\doublespacing

\begin{abstract}
\noindent We analyze the Ben-Porath model of human capital accumulation and several of its variations in order to inform researchers on the consequences of their modeling decisions. We begin with the baseline model which builds on the neutrality assumption, i.e. the assumption that investment and human capital enter in a multiplicative way to the production function of human capital. We explore the properties of the earnings function, the specialization period, the features of the particular case where the production function takes a Cobb-Douglas functional form, and the properties of the returns to schooling. Then we study the Haley-Rosen specification, a finite horizon version with a particular Cobb-Douglas parameterization. We also work on a generalization of the Cobb-Douglas case where neutrality is no longer true, but depreciation is set to zero. Finally, we work on the Shenshinski specification, a general Cobb-Douglas formulation with positive depreciation leading to bang-bang equlibria solutions.
\end{abstract}

\input{Sections/Introduction}
\input{Sections/BasicModel}
\input{Sections/Literature} 

\clearpage
\bibliographystyle{chicago}
\bibliography{BibtexFiles/BenPorath}


\end{document}
