\section{Introduction}
\noindent Human capital investment and accumulation are essential components for many studies on a wide range of topics in Economics, including the economic growth and development of nations, the gender and black-white wage gaps, and the rate of return to schooling. \citet{becker1962investment} initiates the formal analysis of human capital studies and offers the first unified and comprehensive framework to study human capital investment with the standard Economic tools.\footnote{\citet{becker2009human} collates the work of the author on the subject which covers schooling, learning-by-doing, and on-the-job training.} \\
\indent Following Becker's work, \citet{ben1967production} proposes a dynamic model that relates human capital accumulation to life-cycle earnings. Currently, this is the workhorse model when it comes to analyzing the relationship between human capital accumulation decisions and life-cycle outcomes. The vast majority of papers that model human capital use some variation of the Ben-Porath model. These variations lead to different time profiles of investment in human capital, human capital accumulation, and earnings.\\
\indent In this paper, we analyze the Ben-Porath model and several of its variations in order to inform researchers on the consequences of their modeling decisions. We proceed as follows: Section 2 presents the baseline specification. Many implications can be derived in closed form and the intuition can be easily explained with straightforward graphical analysis based on the assumptions of no depreciation and a finite horizon. Section 3 analyzes the Haley-Rosen specification, which allows a finite horizon but still keeps tractability. %You said this version is finite time but you also said the basic model is finite time. Did you mean to say the basic model is infinite time?
 Section 4 studies the model in a more general formulation by relaxing the neutrality assumption. Section 5 studies a specification that generates a Bang-Bang equilibrium. Section 6 allows for depreciation and develops conditions under which investment in human capital happens in different episodes over the life-cycle. Section 7 offers some final comments. \\